\Huge Calibration of the High Momentum Spectrometer Drift Chambers

### 2019-06-21 Fellowships/NRC_report/NRC_report/report.tex
update

\newcommand{\CLASSINPUTtoptextmargin}{2cm}
\newcommand{\CLASSINPUTbottomtextmargin}{1in}
\documentclass[journal, a4paper]{IEEEtran}

@@

\title{\huge Calibration of the High Momentum Spectrometer Drift Chambers}
\author{Carlos Yero\\March 20, 2017}

@@

neede to be satisfy by Hall C to demonstrate that the SHMS was in operating conditions, data was taken with actual beam in the

@@

The calibration run taken had enough statistics, that even at the edge wires had sufficient events for a linear

@@

The weighted average over all wires is assigned as the $``t_{0}"$ of the plane. When this procedure is done for all planes

@@

d_{drift}(\tau = T) = \frac{\Delta}{2} \frac{\int_{t_{0}}^{T<t_{max}}F(\tau)d\tau}{\int_{t_{0}}^{t_{max}}F(\tau)d\tau}

@@

Drift velocities of free electrons are sensitive to the gas mixture used in

@@

During the calibration run, the HMS drift chambers operated at a 50:50 mixture by volume of Argon/Ethane. The drift velocities

@@

The Y'-plane in chamber 2 shows a significant deviation from the rest. This could indicate a possible variation in the

@@

The linear correlation that exists between drift time and distance in Figure \ref{fig:drift_velocity} shows

@@

\section{Drift Chamber Efficiencies and Residuals}
The per plane efficciencies were determined for all planes. The efficinecy of any given plane
in a chamber is defined as the ratio of the number of hits that were detected by the plane to the number of hits that
\textit{should} have been detected by that plane.
Where a \textit{hit} is defined as a detectable signal in the chamber from ionized charges produced by the passage of a
particle. Mathematically, the efficiency of the $i^{th}$ plane is given by \\
\begin{equation*}
\epsilon_{i} = \frac{\# \text{ hits that were detected by the $i^{th}$ plane}}{\# \text{ hits that \textit{should} have been detected by the $i^{th}$ plane}}
\end{equation*}
with the condition that the hit was also detected by the remaining five planes in the chamber. The
efficiencies of both chambers are summarized in Table \ref{tab:eff}.
\begin{table}[!h]
  \scalebox{0.8} {
\begin{tabular}{ |l|l|l| }
\hline
\multicolumn{3}{ |c| }{\textbf{HMS Drift Chamber Efficiencies}} \\
\hline
\textbf{Plane} & \textbf{Chamber 1} (\%) & \textbf{Chamber 2} (\%) \\ \hline
\multirow{6}{*}{}
X  & 95.8 & 99.4  \\ \hline
Y  & 98.7 & 99.7  \\ \hline
U  & 97.7 & 96.6  \\ \hline
V  & 96.5 & 97.7  \\ \hline
Y' & 99.0 & 97.7  \\ \hline
X' & 94.9 & 97.2  \\ \hline
\end{tabular}
}
\caption{Drift chamber plane efficiencies}
\label{tab:eff}
\end{table}\\
\indent The best way to determine the drift chamber performance is by measuring the spatial resolution, or how
well it can measure the position of particle tracks. This measurement is done through the determination of per plane
residuals. For a particle traversing at least 4 planes of the chamber, a collection of space-points (X,Y) is measured based on the
wires that fired. The space points are fitted with a straight line such as to minimize the chi-square, and obtain a best
fit. The line fit is then compared to the meausured space-point from the plane wires that fired, and the difference is called the \textit{residual}
for that plane. The residuals are calculated on an event by event basis, and should be centered around zero. (See Figure \ref{fig:hdc1y1_residual})
\begin{figure}[!ht]
  \centering
  \includegraphics[width=3.6in, height=2.7in]{hdc1y1_residuals.pdf}
  \caption{Residuals for the Y-plane in chamber 1. The Full Width at Half Maximum (FWHM or $\sigma$)  is representative the
  spatial resolution.}
  \label{fig:hdc1y1_residual}
\end{figure}\\
A gaussian fit was done for all planes in both chambers to extract the space resolution of the chambers. The fit results are
summarized in Table \ref{tab:res}
\begin{table}[!h]
\scalebox{0.8} {
\begin{tabular}{ |l|l|l| }
\hline
\multicolumn{3}{ |c| }{\textbf{HMS Drift Chamber Residuals}} \\
\hline
\textbf{Plane} & \textbf{Chamber 1}, $\sigma(\mu m)$ & \textbf{Chamber 2}, $\sigma(\mu m)$ \\ \hline
\multirow{6}{*}{}
X & 416$\pm$1.69 & 388$\pm$1.60 \\ \hline
Y & 360$\pm$1.18 & 408$\pm$1.22 \\ \hline
U & 361$\pm$1.29 & 327$\pm$1.27 \\ \hline
V & 382$\pm$1.47 & 323$\pm$1.28 \\ \hline
Y'& 383$\pm$1.31 & 391$\pm$1.16 \\ \hline
X'& 419$\pm$1.63 & 384$\pm$1.55 \\ \hline
\end{tabular}
}
\caption{Drift chamber plane residuals}
\label{tab:res}
\end{table}

\section{Summary}
This report has outlined the steps taken to calibrate the HMS drift chambers using data taken on March 2017. The outline
also covered some of the difficulties encountered in the calibration procedure, and the steps taken to overcome them. The effects
of the gas mixture and applied high voltage on the drift velocities was also briefly discussed. Finally, two
performance criterions (efficiencies and residuals) used to describe the performance of the drift chambers were investigated. The
efficiencies described how well were the chambers able to detect hits produced by the passage of a particle. The residuals described
how well does the measured particle position in each plane compare to the best track fit through all six planes in each chamber.

>>>>

\newcommand{\CLASSINPUTtoptextmargin}{1.5cm}
\newcommand{\CLASSINPUTbottomtextmargin}{1.5cm}
\documentclass[journal, a4paper]{IEEEtran}

@@

\title{\Huge Calibration of the High Momentum Spectrometer Drift Chambers}
\author{Carlos Yero\\March 22, 2017}

@@

needed to be satisfy by Hall C to demonstrate that the SHMS was in operating conditions, data was taken with actual beam in the

@@

The calibration run taken had enough statistics, that even at the edge, wires had sufficient events for a linear

@@

\indent The weighted average over all wires is assigned as the $``t_{0}"$ of the plane. When this procedure is done for all planes

@@

d_{drift}(\tau = T) = \frac{\Delta}{2} \frac{\int_{t_{0}}^{T\leq t_{max}}F(\tau)d\tau}{\int_{t_{0}}^{t_{max}}F(\tau)d\tau}

@@

\noindent Drift velocities of free electrons are sensitive to the gas mixture used in

@@

\indent During the calibration run, the HMS drift chambers operated at a 50:50 mixture by volume of Argon/Ethane. The drift velocities

@@

\indent The Y'-plane in chamber 2 shows a significant deviation from the rest. This could indicate a possible variation in the

@@

\indent The linear correlation that exists between drift time and distance in Figure \ref{fig:drift_velocity} shows

@@

\section{Drift Chamber Efficiencies and Residuals}
\noindent The per plane efficciencies were determined for all planes. The efficinecy of any given plane
in a chamber is defined as the ratio of the number of hits that were detected by the plane to the number of hits that
\textit{should} have been detected by that plane.
Where a \textit{hit} is defined as a detectable signal in the chamber from ionized charges produced by the passage of a
particle. Mathematically, the efficiency of the $i^{th}$ plane is given by \\
\begin{equation*}
\epsilon_{i} = \frac{\# \text{ hits that were detected by the $i^{th}$ plane}}{\# \text{ hits that \textit{should} have been detected by the $i^{th}$ plane}}
\end{equation*}
with the condition that the hit was also detected by the remaining five planes in the chamber. The
efficiencies of both chambers are summarized in Table \ref{tab:eff}.
\begin{table}[!h]
\scalebox{0.8} {
\begin{tabular}{ |l|l|l| }
\hline
\multicolumn{3}{ |c| }{\textbf{HMS Drift Chamber Efficiencies}} \\
\hline
\textbf{Plane} & \textbf{Chamber 1} (\%) & \textbf{Chamber 2} (\%) \\ \hline
\multirow{6}{*}{}
X  & 95.8 & 99.4  \\ \hline
Y  & 98.7 & 99.7  \\ \hline
U  & 97.7 & 96.6  \\ \hline
V  & 96.5 & 97.7  \\ \hline
Y' & 99.0 & 97.7  \\ \hline
X' & 94.9 & 97.2  \\ \hline
\end{tabular}
}
\caption{Drift chamber plane efficiencies}
\label{tab:eff}
\end{table}\\
\indent The best way to determine the drift chamber performance is by measuring the spatial resolution, or how
well it can measure the position of particle tracks. This measurement is done through the determination of per plane
residuals. For a particle traversing at least 4 planes of the chamber, a collection of space-points (X,Y) is measured based on the
wires that fired. The space points are fitted with a straight line such as to minimize the chi-square, and obtain a best
fit. The line fit is then compared to the meausured space-point from the plane wires that fired, and the difference is called the \textit{residual}
for that plane. The residuals are calculated on an event by event basis, and should be centered around zero. (See Figure \ref{fig:hdc1y1_residual})
  \includegraphics[width=3.6in, height=2.7in]{hdc1y1_residuals.pdf}
  \caption{Residuals for the Y-plane in chamber 1. The Full Width at Half Maximum (FWHM or $\sigma$)  is representative the
  spatial resolution.}
  \label{fig:hdc1y1_residual}
\indent A gaussian fit was done for all planes in both chambers to extract the space resolution of the chambers. The fit results are
summarized in Table \ref{tab:res}
\begin{table}[!h]
\scalebox{0.8} {
\begin{tabular}{ |l|l|l| }
\hline
\multicolumn{3}{ |c| }{\textbf{HMS Drift Chamber Residuals}} \\
\hline
\textbf{Plane} & \textbf{Chamber 1}, $\sigma(\mu m)$ & \textbf{Chamber 2}, $\sigma(\mu m)$ \\ \hline
\multirow{6}{*}{}
X & 416$\pm$1.69 & 388$\pm$1.60 \\ \hline
Y & 360$\pm$1.18 & 408$\pm$1.22 \\ \hline
U & 361$\pm$1.29 & 327$\pm$1.27 \\ \hline
V & 382$\pm$1.47 & 323$\pm$1.28 \\ \hline
Y'& 383$\pm$1.31 & 391$\pm$1.16 \\ \hline
X'& 419$\pm$1.63 & 384$\pm$1.55 \\ \hline
\end{tabular}
}
\caption{Drift chamber plane residuals}
\label{tab:res}
\end{table}\\
The residuals determined in Table \ref{tab:res} include contributions from the residuals of the other five planes in each chamber \cite{baker}. From fitted curve in Figure \ref{fig:hdc1y1_residual}, the resolutions of the remaining five planes were added in quadrature. In general, the resolution of the $i^{th}$ plane in each chamber can determined by the following equation
\sigma^{2}_{i} = \sigma^{2}_{1} +\sigma^{2}_{2} +\sigma^{2}_{3} +\sigma^{2}_{4} +\sigma^{2}_{5} = 5\sigma^{2}  
since it was assumed that the other five planes have equal contributions to the
residual of the sixth plane. 
The per plane residuals can then be determined by dividing the residual determined from the fit ($\sigma_{i}^{2}$) by 5, and taking the square root. For example, the intrinsic resolution of the V-plane in chamber 1 would be $\sigma_{V}=170.8\pm0.657 \mu$m. The intrinsic resolutions per plane are summarized in Table \ref{tab:res2}.
\begin{table}[!h]
\scalebox{0.8} {
\begin{tabular}{ |l|l|l| }
\hline
\multicolumn{3}{ |c| }{\textbf{HMS Drift Chamber Intrinsic Resolutions}} \\
\hline
\textbf{Plane} & \textbf{Chamber 1}, $\sigma(\mu m)$ & \textbf{Chamber 2}, $\sigma(\mu m)$ \\ \hline
\multirow{6}{*}{}
X & 186$\pm$0.755 & 173.5$\pm$0.715 \\ \hline
Y & 160.9$\pm$0.527 & 182.4$\pm$0.545 \\ \hline
U & 161.4$\pm$0.576 & 146.2$\pm$0.567 \\ \hline
V & 170.8$\pm$0.657 & 144.4$\pm$0.572 \\ \hline
Y'& 171.2$\pm$0.585 & 174.8$\pm$0.518 \\ \hline
X'& 187.3$\pm$0.728 & 171.7$\pm$0.693 \\ \hline
\end{tabular}
}
\caption{Drift chamber intrincis resolutions per plane}
\label{tab:res2}
\end{table}\\

\section{Summary}
\noindent This report has outlined the steps taken to calibrate the HMS drift chambers using data taken on March 2017. The outline
also covered some of the difficulties encountered in the calibration procedure, and the steps taken to overcome them. The effects
of the gas mixture and applied high voltage on the drift velocities was also briefly discussed. Finally, two
performance criterions (efficiencies and residuals) used to describe the performance of the drift chambers were investigated. The
efficiencies described how well were the chambers able to detect hits produced by the passage of a particle. The residuals described
how well does the measured particle position in each plane compare to the best track fit through all six planes in each chamber.
